A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2019-01-28 doc/large_strain.tex

In order to have a better understanding of the behavior of the most promising ``tensor4'' implementation,
we collect its performance data and computing times for various stages inside the element loop (see Algorithm \ref{alg:mf_tensor4}):
(i) `read / write' denotes reading and writing from/to a global vector (step 1 and 13);
(ii) `sum factorization' denotes application of sum factorization techniques (steps 4, 5 and 12);
(iii) `quadrature loop' denotes operations performed on each quadrature (steps 6 -- 11).
Given that the measurements are now down within the cell loop, we collected results on smaller meshes by adopting only one
global refinement in 2D and no global mesh refinement in 3D.
It is clear from Figure \ref{fig:breakdown_stackedbar_2d_tensor4} and  \ref{fig:breakdown_stackedbar_3d_tensor4} that
`read / write' steps takes negligible amount of time spent on each element.
For higher order bases the majority of time is spent within the quadrature loop.
When analyzing the Roofline data (Figure \ref{fig:roofline_2d_tensor4} and \ref{fig:roofline_3d_tensor4}),
we observe that this part of the algorithm demonstrates rather low performance, which in our opinion is the source of the
suboptimal performance of the whole matrix-free matrix-vector product (see Figure \ref{fig:roofline}).
In this case the quadrature loop consists of, essentially, a
single contraction between two second order tensors (step 7 in Algorithm \ref{alg:mf_tensor4}) and a
double contraction between a symmetric second order tensor and a fourth order tensor (step 9 in Algorithm \ref{alg:mf_tensor4}).
Further optimization of \texttt{Tensor} classes in \textit{deal.II} are required to
improve the performance.
Such studies, however, are beyond the scope of the current work.
}

>>>>

In order to have a better understanding of the behavior of the most promising ``tensor4'' implementation,
we collect its performance data and computing times for various stages inside the element loop (see Algorithm \ref{alg:mf_tensor4}):
(i) `read / write' denotes reading and writing from/to a global vector (step 1 and 13);
(ii) `sum factorization' denotes application of sum factorization techniques (steps 4, 5 and 12);
(iii) `quadrature loop' denotes operations performed on each quadrature (steps \mbox{6 -- 11}).
Given that the measurements are now down within the cell loop, we collected results on smaller meshes by adopting only one
global refinement in 2D and no global mesh refinement in 3D.
It is clear from Figure \ref{fig:breakdown_stackedbar_2d_tensor4} and  \ref{fig:breakdown_stackedbar_3d_tensor4} that
`read / write' steps takes negligible amount of time spent on each element.
For higher order bases the majority of time is spent within the quadrature loop.
When analyzing the Roofline data (Figure \ref{fig:roofline_2d_tensor4} and \ref{fig:roofline_3d_tensor4}),
we observe that this part of the algorithm demonstrates rather low performance, which in our opinion is the source of the
suboptimal performance of the whole matrix-free matrix-vector product (see Figure \ref{fig:roofline}).
In this case the quadrature loop consists of, essentially, a
single contraction between two second order tensors (step 7 in Algorithm \ref{alg:mf_tensor4}) and a
double contraction between a symmetric second order tensor and a fourth order tensor (step 9 in Algorithm \ref{alg:mf_tensor4}).
Further optimization of \texttt{Tensor} classes in \textit{deal.II} are required to
improve the performance.
Such studies, however, are beyond the scope of the current work.
}
